Meta-analysis of fungicide application timing for powdery mildew management

### 2020-02-24 DataWrangle/ExcludeBook_191115_PMMB_DataWrangling_PM.Rmd
Update ExcludeBook_191115_PMMB_DataWrangling_PM.Rmd

# Data wrangling
This file is not complete and there are a number of things that need checking
   - Mean squares are only calculated for the triazoles and controls
   - If the years for each trial are input correctly
   - If the disease and yield errors are calculated and that they are shuffled correctly into the data frame
   - AUDPC needs to be checked for starting dates and if trials have been omitted
   - Check that what is generated from this file matches the main file and update
   - write.csv should have row.names = FALSE

## Import Libraries
```{r message=FALSE, warning=FALSE}
# install.packages("openxlsx")
# install.packages("readxl")
# install.packages("tidyr")
# install.packages("agricolae")

library(openxlsx)
library(readxl)
library(tidyr)
library(dplyr)
library(data.table)
library(agricolae)

@@

## Functions
```{r functions}

fold_data <- function(new_DataFrame, template_DataFrame){
   for(i in colnames(new_DataFrame)){
      if(i == colnames(new_DataFrame)[1]){
         x1 <- vector()}

@@

      x1 <- c(x1,match(i, cnames2, nomatch = NA)) # MAke a index vector for the location of new_DataFrame col names in template_DataFrame
   }

   new_DataFrame[cnames2[!(cnames2 %in% cnames1)]] <- NA # adding column names to new_DataFrame not in template_DataFrame
   new_DataFrame <- 
      new_DataFrame[,c(cnames2, cnames1[is.na(x1)])]  # adding column names to new_DataFrame not in template_DataFrame
   return(new_DataFrame)
}
source("../copy_class.R")
source("../R/data_mesh.R")
```

@@

Herm_10 <-
   as.data.frame(read_xls(
      "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2010/2010 PMmung Herm/2010 PMmung hermitage RAW field book.xls",
      sheet = "For analysis"
   ), stringsAsFactors == FALSE)

Herm_10 <- Herm_10 %>%
   mutate(trade_name = case_when(`Schedule!` == "U/S" ~ NA_character_,
                                 `Schedule!` == "FirstsignC" ~ " ",
                                 TRUE ~ "Kendon"))%>%
   mutate(fungicide = case_when(`Schedule!` == "U/S" ~ "control",
                                 `Schedule!` == "FirstsignC" ~ "carbendazim",
                                 TRUE ~ "sulphur"))%>%
   mutate(dose_ai.ha = case_when(`Schedule!` == "U/S" ~ 0,
                                 `Schedule!` == "FirstsignC" ~ 250,
                                 TRUE ~ 2400))%>%
   mutate(fungicide_application_1 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
                                              `Schedule!` == "Early Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
                                              `Schedule!` == "FirstsignC" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              `Schedule!` == "FirstsignS" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              `Schedule!` == "U/S" ~ as.Date(NA),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_2 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
                                              `Schedule!` == "Early Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_3 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-05", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_4 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-11", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_5 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_6 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-25", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_7 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-07", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(total_fungicide = case_when(`Schedule!` == "Full Sp" ~ 7,
                                      `Schedule!` == "Early Sp" ~ 2,
                                      `Schedule!` == "FirstsignC" ~ 1,
                                      `Schedule!` == "FirstsignS" ~ 1,
                                      `Schedule!` == "U/S" ~ 0,
                                      TRUE ~ 9999))

@@

```{r Herm10_dataInput}
Herm_10_dat <- 
   data.frame(trial_ref = "mung0910/01",
              year = "2010",
              location = "Hermitage",
              host_genotype = Herm_10$`Cultivar!`,
              trial_design = "RB",
              plot_length = Herm_10$`Length (m)`,
              plot_width = 4,
              plant_density = NA,
              row_spacing = 0.25,
              replicate = Herm_10$`Rep!`,
              planting_date = as.Date("2010-01-22"),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = as.Date("2010-03-18"),
              trade_name = Herm_10$trade_name,
              fungicide_ai = Herm_10$fungicide,
              dose_ai.ha = Herm_10$dose_ai.ha,
              n_treatment = Herm_10$`Treat!`,
              fungicide_application_1 = Herm_10$fungicide_application_1,
              fungicide_application_2 = Herm_10$fungicide_application_2,
              fungicide_application_3 = Herm_10$fungicide_application_3,
              fungicide_application_4 = Herm_10$fungicide_application_4,
              fungicide_application_5 = Herm_10$fungicide_application_5,
              fungicide_application_6 = Herm_10$fungicide_application_6,
              fungicide_application_7 = Herm_10$fungicide_application_7,
              total_fungicide = Herm_10$total_fungicide,
              harvest_date = as.Date("2010-05-14"),
              final_assessment = as.Date("2010-04-20"),
              PM_final_severity = Herm_10$P.M.13wap,
              rating_scale = "1-9",
              grain_yield.t.ha. = Herm_10$`Kg/ha`/1000,
              stringsAsFactors = FALSE
              )

@@

Herm_10_dat <- cbind(Herm_10_dat,
                     AUDPC = apply(data.frame(zero = 0,
                                              Herm_10[,c("P.M.9wap","P.M.11wap","P.M.13wap")]-1),
                                   1, 
                                   FUN = audpc,
                                   dates = c((Herm_10_dat$first_sign_disease[1]-7)-
                                                Herm_10_dat$planting_date[1],
                                             9*7,11*7,13*7)
                                   ),
                     AUDPS = apply(data.frame(zero = 0,
                                                   Herm_10[,c("P.M.9wap","P.M.11wap","P.M.13wap")]-1),
                                        1, 
                                        FUN = audps, 
                                        dates = c((Herm_10_dat$first_sign_disease[1]-7)-
                                                     Herm_10_dat$planting_date[1],
                                                  9*7,11*7,13*7)
                                        )
                          )

Herm_10$PM_zero <- 1
Herm_plot <- melt(Herm_10, id.vars = c("Description!","Treat!", "Rep!","Cultivar!","fungicide","total_fungicide"),
                  measure.vars = c("PM_zero","P.M.9wap","P.M.11wap","P.M.13wap"))

Herm_plot <- Herm_plot %>%
   mutate(T_var = case_when(variable == "PM_zero" ~ 
                               as.numeric((Herm_10_dat$first_sign_disease[1]-7) - Herm_10_dat$planting_date[1]),
                            variable == "P.M.9wap" ~ 9*7,
                            variable == "P.M.11wap" ~ 11*7,
                            variable == "P.M.13wap" ~ 13 * 7)
   )
Herm_plot$`Treat!` <- as.factor(Herm_plot$`Treat!`)
Herm_plot %>%
   ggplot(aes(x = T_var, y = value, color = `Treat!`, shape = `Treat!`))+
   geom_jitter(width = 1)+
   geom_smooth()

@@

Herm_10_m <- Herm_10_dat %>%
   group_by(trial_ref, year, location, host_genotype, trial_design, trade_name,fungicide_ai,dose_ai.ha,n_treatment) %>%
   summarise(plot_length.m. = mean(plot_length, na.rm = TRUE),
              plot_width.m. = mean(plot_width, na.rm = TRUE),
              plant_density = unique(plant_density),
              row_spacing = unique(row_spacing),
              replicates = length(replicate),
              planting_date = unique(planting_date),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = unique(first_sign_disease),
              fungicide_application_1 = unique(fungicide_application_1),
              fungicide_application_2 = unique(fungicide_application_2),
              fungicide_application_3 = unique(fungicide_application_3),
              fungicide_application_4 = unique(fungicide_application_4),
              fungicide_application_5 = unique(fungicide_application_5),
             fungicide_application_6 = unique(fungicide_application_6),
             fungicide_application_6 = unique(fungicide_application_7),
              total_fungicide = unique(total_fungicide),
              harvest_date = unique(harvest_date),
              final_assessment = unique(final_assessment),
              rating_scale = unique(rating_scale),
              disease_error = sd(PM_final_severity, na.rm = TRUE),
              PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
              D_error_type = "stdev",
              raw_graded = NA,
              yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
              grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
              Y_error_type = "stdev",
              AUDPC_m = mean(AUDPC),
             AUDPC_sd = sd(AUDPC),
             AUDPS_m = mean(AUDPS),
             AUDPS_sd = sd(AUDPS),
             Inc_Ms = anova(lm(AUDPC ~ host_genotype + factor(n_treatment), data = Herm_10_dat))[2,3],
             Y_Msquare = anova(lm(grain_yield.t.ha. ~ host_genotype + factor(n_treatment), data = Herm_10_dat))[2,3]

@@

Herm_tentmp <- Herm_10_dat %>% 
   group_by(host_genotype,n_treatment)%>%
   summarise(disease_error = sd(PM_final_severity, na.rm = TRUE),
             yield_error = sd(grain_yield.t.ha., na.rm = TRUE))

Herm_10_m[order(Herm_10_m$n_treatment),"disease_error"] <- Herm_tentmp[order(Herm_tentmp$n_treatment),"disease_error"]
Herm_10_m[order(Herm_10_m$n_treatment),"yield_error"] <- Herm_tentmp[order(Herm_tentmp$n_treatment),"yield_error"]

@@

source("../R/data_mesh.R")
# insert new calculations into data
PM_MB_means <- replace_data(template_data = PM_MB_means, 
                            new_data = Herm_10_m, 
                            new_data_column = c("AUDPC_m", "AUDPC_sd", "AUDPS_m","AUDPS_sd","Y_Msquare", "Inc_Ms"),
                            filter_cols = c("trial_ref", "year", "location", "host_genotype", 
                                           "fungicide_ai", "dose_ai.ha","n_treatment"))
                            #ignore_columns = c(6,17,30:ncol(Herm_10_m))

write.csv(
   as.data.frame(Herm_10_m),
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2010/2010 PMmung Herm/2010 PMmung Hermitage means.csv"
)

>>>>

# Data wrangling
This file is not complete and there are a number of things that need checking

   - Mean squares are only calculated for the triazoles and controls

   - If the years for each trial are input correctly

   - If the disease and yield errors are calculated and that they are shuffled correctly into the data frame

   - AUDPC needs to be checked for starting dates and if trials have been omitted

   - Check that what is generated from this file matches the main file and update

   - write.csv should have row.names = FALSE

## Import Libraries
```{r message=FALSE, warning=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               readxl,
               data.table,
               devtools,
               agricolae)
source("R/import_data.R")
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
#source("./getPPD2.R")
theme_set(theme_usq()
)

@@

## Functions
```{r functions}
fold_data <- function(new_DataFrame, template_DataFrame) {
   for (i in colnames(new_DataFrame)) {
      if (i == colnames(new_DataFrame)[1]) {
         x1 <- vector()
      }

@@

      x1 <-
         c(x1, match(i, cnames2, nomatch = NA)) # MAke a index vector for the location of new_DataFrame col names in template_DataFrame
   }

   new_DataFrame[cnames2[!(cnames2 %in% cnames1)]] <-
      NA # adding column names to new_DataFrame not in template_DataFrame
   new_DataFrame <-
      new_DataFrame[, c(cnames2, cnames1[is.na(x1)])]  # adding column names to new_DataFrame not in template_DataFrame
   return(new_DataFrame)
}
source("../copy_class.R")
source("../R/data_mesh.R")
```

@@

Herm_10 <-
   as.data.frame(
      read_xls(
         "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2010/2010 PMmung Herm/2010 PMmung hermitage RAW field book.xls",
         sheet = "For analysis"
      ),
      stringsAsFactors == FALSE
   )

Herm_10 <- Herm_10 %>%
   mutate(
      trade_name = case_when(
         `Schedule!` == "U/S" ~ NA_character_,
         `Schedule!` == "FirstsignC" ~ " ",
         TRUE ~ "Kendon"
      )
   ) %>%
   mutate(
      fungicide = case_when(
         `Schedule!` == "U/S" ~ "control",
         `Schedule!` == "FirstsignC" ~ "carbendazim",
         TRUE ~ "sulphur"
      )
   ) %>%
   mutate(dose_ai.ha = case_when(
      `Schedule!` == "U/S" ~ 0,
      `Schedule!` == "FirstsignC" ~ 250,
      TRUE ~ 2400
   )) %>%
   mutate(
      fungicide_application_1 = case_when(
         `Schedule!` == "Full Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
         `Schedule!` == "Early Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
         `Schedule!` == "FirstsignC" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
         `Schedule!` == "FirstsignS" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
         `Schedule!` == "U/S" ~ as.Date(NA),
         TRUE ~ as.Date(NA)
      )
   ) %>%
   mutate(
      fungicide_application_2 = case_when(
         `Schedule!` == "Full Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
         `Schedule!` == "Early Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
         TRUE ~ as.Date(NA)
      )
   ) %>%
   mutate(fungicide_application_3 = case_when(
      `Schedule!` == "Full Sp" ~ as.Date("2010-03-05", format = "%Y-%m-%d"),
      TRUE ~ as.Date(NA)
   )) %>%
   mutate(fungicide_application_4 = case_when(
      `Schedule!` == "Full Sp" ~ as.Date("2010-03-11", format = "%Y-%m-%d"),
      TRUE ~ as.Date(NA)
   )) %>%
   mutate(fungicide_application_5 = case_when(
      `Schedule!` == "Full Sp" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
      TRUE ~ as.Date(NA)
   )) %>%
   mutate(fungicide_application_6 = case_when(
      `Schedule!` == "Full Sp" ~ as.Date("2010-03-25", format = "%Y-%m-%d"),
      TRUE ~ as.Date(NA)
   )) %>%
   mutate(fungicide_application_7 = case_when(
      `Schedule!` == "Full Sp" ~ as.Date("2010-03-07", format = "%Y-%m-%d"),
      TRUE ~ as.Date(NA)
   )) %>%
   mutate(
      total_fungicide = case_when(
         `Schedule!` == "Full Sp" ~ 7,
         `Schedule!` == "Early Sp" ~ 2,
         `Schedule!` == "FirstsignC" ~ 1,
         `Schedule!` == "FirstsignS" ~ 1,
         `Schedule!` == "U/S" ~ 0,
         TRUE ~ 9999
      )
   )

@@

```{r Herm10_dataInput}
Herm_10_dat <- 
   data.frame(
      trial_ref = "mung0910/01",
      year = "2010",
      location = "Hermitage",
      host_genotype = Herm_10$`Cultivar!`,
      trial_design = "RB",
      plot_length = Herm_10$`Length (m)`,
      plot_width = 4,
      plant_density = NA,
      row_spacing = 0.25,
      replicate = Herm_10$`Rep!`,
      planting_date = as.Date("2010-01-22"),
      emergence_date = NA,
      flowering_date = NA,
      pod_fill_date = NA,
      mid_late_pod_fill = NA,
      first_sign_disease = as.Date("2010-03-18"),
      trade_name = Herm_10$trade_name,
      fungicide_ai = Herm_10$fungicide,
      dose_ai.ha = Herm_10$dose_ai.ha,
      n_treatment = Herm_10$`Treat!`,
      fungicide_application_1 = Herm_10$fungicide_application_1,
      fungicide_application_2 = Herm_10$fungicide_application_2,
      fungicide_application_3 = Herm_10$fungicide_application_3,
      fungicide_application_4 = Herm_10$fungicide_application_4,
      fungicide_application_5 = Herm_10$fungicide_application_5,
      fungicide_application_6 = Herm_10$fungicide_application_6,
      fungicide_application_7 = Herm_10$fungicide_application_7,
      total_fungicide = Herm_10$total_fungicide,
      harvest_date = as.Date("2010-05-14"),
      final_assessment = as.Date("2010-04-20"),
      PM_final_severity = Herm_10$P.M.13wap,
      rating_scale = "1-9",
      grain_yield.t.ha. = Herm_10$`Kg/ha` / 1000,
      stringsAsFactors = FALSE
   )

@@

Herm_10_dat <- cbind(
   Herm_10_dat,
   AUDPC = apply(
      data.frame(zero = 0,
                 Herm_10[, c("P.M.9wap", "P.M.11wap", "P.M.13wap")] -
                    1),
      1,
      FUN = audpc,
      dates = c((Herm_10_dat$first_sign_disease[1] -
                    7) -
                   Herm_10_dat$planting_date[1],
                9 * 7,
                11 * 7,
                13 * 7
      )
   ),
   AUDPS = apply(
      data.frame(zero = 0,
                 Herm_10[, c("P.M.9wap", "P.M.11wap", "P.M.13wap")] -
                    1),
      1,
      FUN = audps,
      dates = c((Herm_10_dat$first_sign_disease[1] -
                    7) -
                   Herm_10_dat$planting_date[1],
                9 * 7,
                11 * 7,
                13 * 7
      )
   )
)

Herm_10$PM_zero <- 1
Herm_plot <-
   melt(
      Herm_10,
      id.vars = c(
         "Description!",
         "Treat!",
         "Rep!",
         "Cultivar!",
         "fungicide",
         "total_fungicide"
      ),
      measure.vars = c("PM_zero", "P.M.9wap", "P.M.11wap", "P.M.13wap")
   )

Herm_plot <- Herm_plot %>%
   mutate(
      T_var = case_when(
         variable == "PM_zero" ~
            as.numeric((Herm_10_dat$first_sign_disease[1] -
                           7) - Herm_10_dat$planting_date[1]
            ),
         variable == "P.M.9wap" ~ 9 * 7,
         variable == "P.M.11wap" ~ 11 * 7,
         variable == "P.M.13wap" ~ 13 * 7
      )
   )
Herm_plot$`Treat!` <- as.factor(Herm_plot$`Treat!`)
Herm_plot %>%
   ggplot(aes(
      x = T_var,
      y = value,
      color = `Treat!`,
      shape = `Treat!`
   )) +
   geom_jitter(width = 1) +
   geom_smooth()

@@

Herm_10_m <- Herm_10_dat %>%
   group_by(
      trial_ref,
      year,
      location,
      host_genotype,
      trial_design,
      trade_name,
      fungicide_ai,
      dose_ai.ha,
      n_treatment
   ) %>%
   summarise(
      plot_length.m. = mean(plot_length, na.rm = TRUE),
      plot_width.m. = mean(plot_width, na.rm = TRUE),
      plant_density = unique(plant_density),
      row_spacing = unique(row_spacing),
      replicates = length(replicate),
      planting_date = unique(planting_date),
      emergence_date = NA,
      flowering_date = NA,
      pod_fill_date = NA,
      mid_late_pod_fill = NA,
      first_sign_disease = unique(first_sign_disease),
      fungicide_application_1 = unique(fungicide_application_1),
      fungicide_application_2 = unique(fungicide_application_2),
      fungicide_application_3 = unique(fungicide_application_3),
      fungicide_application_4 = unique(fungicide_application_4),
      fungicide_application_5 = unique(fungicide_application_5),
      fungicide_application_6 = unique(fungicide_application_6),
      fungicide_application_6 = unique(fungicide_application_7),
      total_fungicide = unique(total_fungicide),
      harvest_date = unique(harvest_date),
      final_assessment = unique(final_assessment),
      rating_scale = unique(rating_scale),
      disease_error = sd(PM_final_severity, na.rm = TRUE),
      PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
      D_error_type = "stdev",
      raw_graded = NA,
      yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
      grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
      Y_error_type = "stdev",
      AUDPC_m = mean(AUDPC),
      AUDPC_sd = sd(AUDPC),
      AUDPS_m = mean(AUDPS),
      AUDPS_sd = sd(AUDPS),
      Inc_Ms = anova(lm(
         AUDPC ~ host_genotype + factor(n_treatment), data = Herm_10_dat
      ))[2, 3],
      Y_Msquare = anova(
         lm(grain_yield.t.ha. ~ host_genotype + factor(n_treatment), data = Herm_10_dat)
      )[2, 3]

@@

Herm_tentmp <- Herm_10_dat %>%
   group_by(host_genotype, n_treatment) %>%
   summarise(
      disease_error = sd(PM_final_severity, na.rm = TRUE),
      yield_error = sd(grain_yield.t.ha., na.rm = TRUE)
   )

Herm_10_m[order(Herm_10_m$n_treatment), "disease_error"] <-
   Herm_tentmp[order(Herm_tentmp$n_treatment), "disease_error"]
Herm_10_m[order(Herm_10_m$n_treatment), "yield_error"] <-
   Herm_tentmp[order(Herm_tentmp$n_treatment), "yield_error"]

@@

source("../R/data_mesh.R")
# insert new calculations into data
PM_MB_means <- replace_data(
   template_data = PM_MB_means,
   new_data = Herm_10_m,
   new_data_column = c(
      "AUDPC_m",
      "AUDPC_sd",
      "AUDPS_m",
      "AUDPS_sd",
      "Y_Msquare",
      "Inc_Ms"
   ),
   filter_cols = c(
      "trial_ref",
      "year",
      "location",
      "host_genotype",
      "fungicide_ai",
      "dose_ai.ha",
      "n_treatment"
   )
)
#ignore_columns = c(6,17,30:ncol(Herm_10_m))

write.csv(
   as.data.frame(Herm_10_m),
   "../data/Past trials/2010/2010 PMmung Herm/2010 PMmung Hermitage means.csv"
)
